---
title: "DATA607_Assignment4"
author: "Joyce Aldrich"
date: "2022-10-05"
output:
  html_document: default
  pdf_document: default
---

#1.load csv file into r
```{r}
flight <- read.csv (file ='/Users/joycealdrich/Documents/SPS Data Science/Data 607/Assignment_4/Flight.csv')
head(flight)
```

```{r}
library(tidyr)
library(dplyr)
```

#2 remove empty row
```{r}
flight2 <- flight [-c(3),]
flight2
```
#3 Put the airline name in the x col
```{r}
flight2$X <- c("ALASKA","ALASKA","AM WEST", "AM WEST")
```

#4 organize the table
```{r}
flight2 <- gather(flight2, "Destination", "Count", 3:7)
```
#5 rename the 1st and 2nd col
```{r}
flight2 <- rename(flight2, Airline=X, Status=X.1)
```

#6 reorder the col
```{r}
flight2 <- select(flight2, Airline, Destination, Status,Count)
```

#6 cast status and count
```{r}
flight3 <- flight2 %>%
  pivot_wider(names_from = "Status", values_from = "Count")
head(flight3)
```
#7 adding total col in the data.frame
```{r}
flight4 <- flight3 %>%
  mutate(total=flight3$`on time`+flight3$delayed)
```
#8 adding delayed_rate col in the data.frame 
#Noted that AM WEST airline's delayed rate in 5 different destination all higer than ALASKA airline.
```{r}
flight5 <- flight4 %>%
  mutate(delayed_rate = flight4$delayed/flight4$total)
flight5
```

#9 creating ggplot to compare delayed rate between two airlines and destination
```{r}
library(tidyverse)
flight5 %>%
  ggplot(aes(Airline,delayed_rate,color=Destination))+
  geom_point()
```

